---
title: "SM ES2"
author: "Juliette Limozin"
date: "2023-11-10"
output: html_document
---

```{r setup, include=FALSE}
library(MASS)
```

Check for outliers:

```{r cars}
?hills
pairs(hills)
```

Knock Hill is the outlier? Maybe somne other ones?

```{r pressure, echo=FALSE}
plot(hills$time ~ hills$dist)
plot(hills$time ~ hills$climb)
```


```{r}
hills[hills$time > 50 & hills$dist <5,]
hills[hills$time > 180 & hills$dist > 15,]
```


Knock Hill is our outlier: weirdly long time considering the distance and climb. Bens of Jura seemed like an outlier, but the long time is actually explained by the very high climb over a short distance.


```{r}
hills['Knock Hill', 'time'] <- hills['Knock Hill', 'time'] - 60
```


Data is right skewed (i.e. bunching towards the left). This could mean the fit of a model could be biaseCheck log transformation 

```{r}
plot(log(hills$time) ~ log(hills$dist))
plot(log(hills$time) ~ log(hills$climb))
```

Looking more even spread now.
```{r}
lm1 <- lm(log(time) ~ log(dist), data = hills)
summary(lm1)
plot(lm1)
```
```{r}
lm2 <- lm(log(time) ~ log(dist) + log(climb), data = hills)
summary(lm2)
plot(lm2)
```

```{r}
anova(lm1, lm2)
```

```{r}
predict.lm(lm2, data.frame(dist = 5.3,climb = 1100), interval = 'prediction', level = 0.95)
predict.lm(lm1, data.frame(dist = 5.3,climb = 1100), interval = 'prediction', level = 0.95)
```

```{r}
paste0('Model with climb: ',exp(3.556545),' (', exp(3.224142), ', ',exp(3.888948), ')' )
paste0('Model without climb: ',exp(3.561158),' (', exp(3.11599), ', ',exp(4.006326), ')' )
```


```{r}
attach(cabbages)
boxcox(lm(HeadWt ~ Date + Cult))
plot(HeadWt ~ Date + Cult)
```
```{r}
lm1 <- lm(log(HeadWt) ~ Date + Cult)
lm2 <- lm(log(HeadWt) ~ Cult)
anova(lm2, lm1)
```

Lm1 : $log(Y_{i,j,k}) = \mu  + \alpha_i + \beta_j + \epsilon_{i,j,k}, \space \epsilon_{i,j,k} \sim N(0,\sigma^2)$ where $Y_{i,j,k}$ is headweight of kth cabbaged planted at ith date of jth cultivar.

```{r}
lm1 <- lm(log(HeadWt) ~ Date + Cult + Date*Cult)
lm2 <- lm(log(HeadWt) ~ Date + Cult)
anova(lm2, lm1)
summary(lm1)
```


